---
title: "Supplementary Materials"
subtitle: "Local and Regional Disturbance Around Urban Stormwater Ponds Determines their use as Butterfly Habitat"
author: "Jesse D. Fortier*, Isabella C. Richmond, Mary Ann C. Perron, Frances Pick, and Heather Kharouba"
format: 
  html:
    embed-resources: true
---

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| include: false
#| label: setup


source('script/0-Packages.R')

plant <- read.csv('output/PlantCleanbySite.csv') %>% 
  mutate(Pond = str_remove(Pond, "SWF-"),
         Pond = as.numeric(Pond))
anthro <- st_read("output/AnthroFull.gpkg")
niche <- read.csv('input/ButterflyNicheBreadth.csv')
butt <- read.csv('output/ButterflyCleanbySite.csv')

an <- left_join(niche, anthro, by = "Pond")
anp <- left_join(an, plant, by = "Pond")
ab <- inner_join(anthro, butt, by = join_by("Pond" == "SWP"))
abp <- inner_join(ab, plant, by = "Pond")

```

(\*) University of Ottawa, Department of Earth and Environmental Sciences, 150, Louis-Pasteur Private, Ottawa, Ontario, Canada, K1N 6N5\
email: [jfort073\@uottawa.ca](mailto:jfort073@uottawa.ca){.email}

This supplementary information file contains (1) variable descriptions, (2) descriptive statistics for each study site, (3) summary tables of each model. All associated code can be found at the paper's associated GitHub repository, [here](https://github.com/icrichmond/UrbanButterflies).

------------------------------------------------------------------------

## Variables

**Table S1.** Variable names, acronym (used in R code), and descriptions.

```{r}
#| echo: false
#| eval: true
#| label: variables

var <- read.csv("input/variables.csv")

knitr::kable(var)
```

------------------------------------------------------------------------

## Descriptive Statistics

**Table S2.** Descriptive statistics of butterfly abundance, species richness, and Shannon diversity, and plant species richness and % bloom for all flowering plants and native flowering plants. Variable acronyms/shortforms can be found in Table S1.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: descriptive-stats

plant <- read.csv("output/PlantCleanbySite.csv")
plant <- plant %>% 
  select(nspecies, nnative, avgbloom, avgnatbloom)

butt <- read.csv("output/ButterflyCleanbySite.csv")
butt <- butt %>% 
  select(SWP, abund, SpeciesRichness, Shannon) %>%
  rename(Abund = abund,
         Richness = SpeciesRichness)

df <- cbind(butt, plant)

knitr::kable(df, digits = 2)
```

------------------------------------------------------------------------

## Summary Tables

Table S3. Model summary tables for models concerning anthropogenic land use. The first tab contains the results of the model with butterfly abundance as a response, the second tab with butterfly species richness, and the third tab Shannon diversity. Incidence rate ratios are reported, where decimal numbers indicate negative relationships and numbers 1 and above indicate positive relationships. 'anthroper_400' represents the percentage of anthropogenic land use surrounding the pond within a 400 m buffer. Niche breadth categories are presented in comparison to the intercept, which is the 'non-wetland' niche for this table. 'x' in the table indicates an interaction effect. CI columns are 95% confidence intervals.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: urb-400

urb_ab_400 <- readRDS('large/UrbAbund_400.rds')
urb_n_sr_400 <- readRDS('large/UrbNicheSR_400.rds')
urb_sh_400 <- readRDS('large/UrbShann_400.rds')

tab_model(urb_ab_400, urb_n_sr_400, urb_sh_400, dv.labels = c("Abundance", "Species Richness", "Shannon Diversity"),
          show.r2 = F)

```

Table S4. Model summary tables for models concerning anthropogenic land use. The first tab contains the results of the model with butterfly abundance as a response, the second tab with butterfly species richness, and the third tab Shannon diversity. Incidence rate ratios are reported, where decimal numbers indicate negative relationships and numbers 1 and above indicate positive relationships. 'nspecies' represents the number of flowering species present. 'avgbloom' represents the average percent cover of flowering species at each pond. Niche breadth categories are presented in comparison to the intercept, which is the 'non-wetland' niche for this table. 'x' in the table indicates an interaction effect. CI columns are 95% confidence intervals.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: total-plants

tot_ab <- readRDS('large/TotAbund.rds')
tot_n_sr <- readRDS('large/TotNicheSR.rds')
tot_sh <- readRDS('large/TotShann.rds')

tab_model(tot_ab, tot_n_sr, tot_sh, dv.labels = c("Abundance", "Species Richness", "Shannon Diversity"),
          show.r2 = F)

```

Table S5. Model summary tables for models concerning anthropogenic land use. The first tab contains the results of the model with butterfly abundance as a response, the second tab with butterfly species richness, and the third tab Shannon diversity. Incidence rate ratios are reported, where decimal numbers indicate negative relationships and numbers 1 and above indicate positive relationships. 'nnative' represents the number of native flowering species present. 'avgnatbloom' represents the average percent cover of native flowering species at each pond. Niche breadth categories are presented in comparison to the intercept, which is the 'non-wetland' niche for this table. 'x' in the table indicates an interaction effect. CI columns are 95% confidence intervals.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: native-plants

nat_ab <- readRDS('large/NatAbund.rds')
nat_n_sr <- readRDS('large/NatNicheSR.rds')
nat_sh <- readRDS('large/NatShann.rds')

tab_model(nat_ab, nat_n_sr, nat_sh, dv.labels = c("Abundance", "Species Richness", "Shannon Diversity"),
          show.r2 = F)

```
