---
title: "Supplementary Materials"
subtitle: "Local and Regional Disturbance Around Urban Stormwater Ponds Determines their use as Butterfly Habitat"
author: "Jesse D. Fortier*, Isabella C. Richmond, Mary Ann C. Perron, Frances Pick, and Heather Kharouba"
output: pdf_document
---

(\*) University of Ottawa, Department of Earth and Environmental Sciences, 150, Louis-Pasteur Private, Ottawa, Ontario, Canada, K1N 6N5\
email: [jfort073\@uottawa.ca](mailto:jfort073@uottawa.ca){.email}

This supplementary information file contains (1) variable descriptions, (2) descriptive statistics for each study site, (3) AIC tables ranking the performance of each buffer distance for butterfly abundance, species richness, and Shannon diversity (4) AIC tables ranking different models at each of the top performing buffer distances for butterfly abundance, species richness, and Shannon diversity, and (5) summary tables for the top ranking model, disturbance in all cases, at the top performing buffer distance for butterfly abundance, species richness, and Shannon diversity. All associated code can be found at the paper's associated GitHub repository, [here](https://github.com/icrichmond/UrbanButterflies).

------------------------------------------------------------------------

## Variables

**Table S1.** Variable names, acronym (used in R code), and descriptions.

```{r, echo = F, eval=T}
var <- read.csv("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/input/variables.csv")

knitr::kable(var)
```

------------------------------------------------------------------------

## Descriptive Statistics

**Table S2.** Descriptive statistics of disturbance, butterfly abundance, species richness, and Shannon diversity, and plant species richness and % bloom for all flowering plants and native flowering plants. Variable acronyms/shortforms can be found in Table S1.
```{r, echo = F, eval=T, warning=F, message=F}
library(dplyr)

plant <- read.csv("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/output/PlantCleanbySite.csv")
plant <- plant %>% 
  select(nspecies, nnative, avgbloom, avgnatbloom)

dist <- read.csv("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/input/SamplingArea.csv")
dist$PerCut[is.na(dist$PerCut)] <- 0 
dist <- dist %>% 
  mutate(Disturbance = if_else(PerCut >= 30, "Disturbed", "Undisturbed"))
Dist <- dist$Disturbance

butt <- read.csv("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/output/ButterflyCleanbySite.csv")
butt <- butt %>% 
  select(SWP, abund, SpeciesRichness, Shannon) %>%
  rename(Abund = abund,
         Richness = SpeciesRichness)

df <- cbind(plant, Dist, butt)
df <- df[, c(6, 5, 7:9, 1:4)]

knitr::kable(df, digits = 2)
```


------------------------------------------------------------------------

## Model Descriptions

Due to correlations between several of our explanatory variables, we opted to create three models that tested separate hypotheses: disturbance model, native plant model, and total plant model. Dependent variables include abundance, species richness, and Shannon diversity. Species richness and Shannon diversity were rarified using the `iNEXT` package to account for differences in sampling effort across sites.

The disturbance model includes a binary variable of disturbed/undisturbed, measured by Jesse in the field and is based on % mowed with sites that have over approximately 30% mowed marked as disturbed. We also created buffers around each site and calculated the % area that is anthropogenic in each buffer (anthropogenic defined as settlement or roads). Eight buffer distances were tested: 20 m, 50 m, 100 m, 200 m, 500 m, 1000 m, 2000 m, and 5000 m based on Rivest & Kharouba (2021, *Canadian Journal of Zoology*) and Kuussaari et al. (2021, *Urban Ecosystems*).

The native plants model includes number of native species found at each site and average cover of native plant bloom in each quadrat at each site. The total plant model used the same measures but included all flowering plants observed at each site.

Below represents the general code used for each model:

```{r, echo = T, eval = F}
# disturbance model
dist <- lm(response ~ Disturbance * anthroper, data = df)
# native species model 
nat <- lm(response ~ nnative * avgnatbloom, data = df)
# full species model 
full <- lm(response ~ nspecies * avgbloom, data = df)
# null 
null <- lm(response ~ 1, data = x)
```

------------------------------------------------------------------------

## Buffer Distances

To test which buffer distance performed best, we took the top performing model (disturbance for all) for each response variable and did AICc model ranking at each buffer distance. Species richness and Shannon diversity were best explained at similar scales, around 1 km. Abundance had top-ranking models with much smaller buffer distances.

**Table S3.** AICc table of buffer distances for abundance models. All models compared are disturbance models, as disturbance was the top ranking model for abundance. Numbers in model name refer to the buffer distance (i.e., Modnames = AbDist50 = abundance disturbance model at 50 m buffer distance)

```{r, echo = F, eval=T}
a <- readRDS("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/large/AbundanceBufferAIC.rds")

knitr::kable(a) 
```

**Table S4.** AICc table of buffer distances for species richness models. All models compared are disturbance models, as disturbance was the top ranking model for species richness Numbers in model name refer to the buffer distance (i.e., Modnames = SRDist1000 = species richness disturbance model at 1000 m buffer distance)

```{r, echo = F, eval=T}
sr <- readRDS("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/large/SpeciesRichnessBufferAIC.rds")

knitr::kable(sr) 
```

**Table S5.** AICc table of buffer distances for Shannon diversity models. All models compared are disturbance models, as disturbance was the top ranking model for Shannon diversity. Numbers in model name refer to the buffer distance (i.e., Modnames = ShDist1000 = Shannon diversity disturbance model at 1000 m buffer distance)

```{r, echo = F, eval=T}
sr <- readRDS("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/large/ShannonBufferAIC.rds")

knitr::kable(sr) 
```

------------------------------------------------------------------------

## Top Models

To demonstrate the top model at each of the top buffer distances, here are the AICc tables at the best buffer distances for each response variables. We have AICc tables at each buffer distance but are only presenting them for top buffer distances here as to simplify the information being presented.

### Abundance

**Table S6a.** AICc table ranking abundance models at 50 m, one of three buffer distances found within 2 delta AICc.

```{r, echo = F, eval=T}
a <- readRDS("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/large/AbundanceAIC.rds")

knitr::kable(a$AbDist50) 
```

**Table S6b.** AICc table ranking abundance models at 20 m, one of three buffer distances found within 2 delta AICc.

```{r, echo = F, eval=T}
a <- readRDS("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/large/AbundanceAIC.rds")

knitr::kable(a$AbDist20) 
```

**Table S6c.** AICc table ranking abundance models at 5000 m, one of three buffer distances found within 2 delta AICc.

```{r, echo = F, eval=T}
a <- readRDS("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/large/AbundanceAIC.rds")

knitr::kable(a$AbDist5000) 
```

### Species Richness

**Table S7a.** AICc table ranking species richness models at 1000 m, one of three buffer distances found within 2 delta AICc.

```{r, echo = F, eval=T}
sr <- readRDS("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/large/SpeciesRichnessAIC.rds")

knitr::kable(sr$SRDist1000) 
```

**Table S7b.** AICc table ranking species richness models at 2000 m, one of three buffer distances found within 2 delta AICc.

```{r, echo = F, eval=T}
sr <- readRDS("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/large/SpeciesRichnessAIC.rds")

knitr::kable(sr$SRDist2000) 
```

**Table S7c.** AICc table ranking species richness models at 100 m, one of three buffer distances found within 2 delta AICc.

```{r, echo = F, eval=T}
sr <- readRDS("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/large/SpeciesRichnessAIC.rds")

knitr::kable(sr$SRDist100) 
```

### Shannon Diversity

**Table S8a.** AICc table ranking species richness models at 1000 m, one of two buffer distances found within 2 delta AICc.

```{r, echo = F, eval=T}
sh <- readRDS("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/large/ShannonAIC.rds")

knitr::kable(sh$ShDist1000) 
```

**Table S8b.** AICc table ranking species richness models at 100 m, one of two buffer distances found within 2 delta AICc.

```{r, echo = F, eval=T}
sh <- readRDS("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/large/ShannonAIC.rds")

knitr::kable(sh$ShDist100) 
```

------------------------------------------------------------------------

## Top Model Summaries

We investigate the top models at each buffer distance to see which variables are driving these relationships. An important consideration is that our models are explaining a high amount of variation, even when there are not statistically significant relationships. One factor that I think may be driving this is that we do not actually see very much variation in our dataset which is an important consideration for when we write up the results we have found.

### Abundance

**Table S9a.** Summary table for abundance disturbance model at 50 m buffer distance.

```{r, echo = F, eval=T}
library(magrittr)
ab <- readRDS("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/large/AbundanceDistModels.rds")
ab <- ab$AbDist50 
names(ab$coefficients) <- c('Intercept', 'Undisturbed','Anthropogenic (%)','Interaction')
ab %>% broom::tidy() %>% knitr::kable()
```

**Table S9b.** Summary table for abundance disturbance model at 20 m buffer distance.

```{r, echo = F, eval=T}
library(magrittr)
ab <- readRDS("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/large/AbundanceDistModels.rds")
ab <- ab$AbDist20 
names(ab$coefficients) <- c('Intercept', 'Undisturbed','Anthropogenic (%)','Interaction')
ab %>% broom::tidy() %>% knitr::kable()
```

**Table S9c.** Summary table for abundance disturbance model at 5000 m buffer distance.

```{r, echo = F, eval=T}
library(magrittr)
ab <- readRDS("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/large/AbundanceDistModels.rds")
ab <- ab$AbDist5000 
names(ab$coefficients) <- c('Intercept', 'Undisturbed','Anthropogenic (%)','Interaction')
ab %>% broom::tidy() %>% knitr::kable()
```

### Species Richness

**Table S10a.** Summary table for species richness disturbance model at 1000 m buffer distance.

```{r, echo = F, eval=T}
library(magrittr)
sr <- readRDS("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/large/SpeciesRichnessDistModels.rds")
sr <- sr$SRDist1000 
names(sr$coefficients) <- c('Intercept', 'Undisturbed','Anthropogenic (%)','Interaction')
sr %>% broom::tidy() %>% knitr::kable()
```

**Table S10b.** Summary table for species richness disturbance model at 2000 m buffer distance.

```{r, echo = F, eval=T}
library(magrittr)
sr <- readRDS("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/large/SpeciesRichnessDistModels.rds")
sr <- sr$SRDist2000 
names(sr$coefficients) <- c('Intercept', 'Undisturbed','Anthropogenic (%)','Interaction')
sr %>% broom::tidy() %>% knitr::kable()
```

### Shannon Diversity

**Table S11a.** Summary table for Shannon diversity disturbance model at 1000 m buffer distance.

```{r, echo = F, eval=T}
library(magrittr)
sh <- readRDS("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/large/ShannonDistModels.rds")
sh <- sh$ShDist1000 
names(sh$coefficients) <- c('Intercept', 'Undisturbed','Anthropogenic (%)','Interaction')
sh %>% broom::tidy() %>% knitr::kable()
```

**Table S11b.** Summary table for Shannon diversity disturbance model at 100 m buffer distance.

```{r, echo = F, eval=T}
library(magrittr)
sh <- readRDS("C:/Users/I_RICHMO/Documents/Repositories/UrbanButterflies/large/ShannonDistModels.rds")
sh <- sh$ShDist100 
names(sh$coefficients) <- c('Intercept', 'Undisturbed','Anthropogenic (%)','Interaction')
sh %>% broom::tidy() %>% knitr::kable()
```
